TE_computation

### 2021-01-06 computational model/TE_computation.Rmd
Amending the simulation of the computational data set
Adding more levels of V

# Generating dataframe for computational model
```{r}
#### Set 1: Calculate other variables based on V, W, and B ####
## Creating variables: V, W, and B
# Items on vocabulary measure (e.g. words on the CDI)
V <- 600

@@

## Creating the dataframe
TE_fromVWB <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq) %>%
  # Create other variables
  mutate(D = B*W, # Words produced in the dominant language
         N = (1-B)*W, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= 600)) %>%
  # filter out cases where D or N aren't whole numbers
  filter_at(vars(D, N), any_vars(is.wholenumber(.)))

#### Set 2: Calculate other variables based on V, D, and N ####
## Creating variables: D and N
D <- seq(0, V, by = 25)

N <- seq(0, V, by = 25)

@@

## Creating the dataframe
TE_fromVDN <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq) %>%
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # Create other variables
  mutate(W = D+N, # Total word vocabulary of a child
         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>%
  # Remove cases where B = NA
  filter(!is.na(B))

#### Set 3: Calculate other variables based on V, D, and B ####
## Creating variables: D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

@@

## Creating the dataframe
TE_fromVDB <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq) %>%
  # Create other variables
  mutate(W = D/B, # Total word vocabulary of a child
         N = W-D, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= 600)) %>%
  # filter out cases where N isn't whole number
  filter(is.wholenumber(N))

#### Set 4: Calculate other variables based on V, N, and B ####
## Creating variables: D and N
N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

@@

## Creating the dataframe
TE_fromVNB <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq) %>%
  # Create other variables
  mutate(W = D/(1-B), # Total word vocabulary of a child
         D = W-N, # Words produced in dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= 600)) %>%
  # filter out cases where D isn't whole number
  filter(is.wholenumber(D)) %>%
  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.))) %>%
  filter(is.wholenumber(Nu))

#### Combine the 4 sets ####
TE_computation <- TE_fromVWB %>%
  bind_rows(TE_fromVDN, TE_fromVDB, TE_fromVNB) %>%
  # constrain all variables but B (balance) to be integers 
  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))

# Correlation
```{r}
Correlation <- TE_computation %>%
  select(-V) %>%
  cor_matrix()

mycor <- rcorr(as.matrix(TE_computation), type="pearson")
mycor$r
mycor$P

@@

plot_composition_byW <- TE_computation %>%
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N", "C")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  mutate(bin_W = cut_interval(W, 6,
                     labels = c(200, 400, 600, 800, 1000, 1200))) %>%
  group_by(bin_B, vocab_type, bin_W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ bin_W) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
  labs(x = "Balance", 
       y = "Number of concepts", 
       title = "Vocabulary composition across different vocabulary size")

## Compare different vocabulary measures across balance
```{r}
plot_words <- TE_computation %>%
  # Convert to long data
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = as.factor(vocab_type)) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  #mutate(bin_W = cut_interval(W, 12,
  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
  group_by(bin_B, vocab_type, W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = W, y = n_words, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Total vocabulary Size (W)", 
       y = "Number of words", 
       color = "Balance",
       title = "Word vocabulary size by balance under different vocabulary measures") +
  guides(colour = guide_legend(nrow = 1))

>>>>

# Generating dataframe for computational model
#### Set 1: Calculate other variables based on V, W, and B ####
```{r}
### Generating variables: V, W, and B
## V = 600 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 600

@@

# Creating the dataframe
TE_V600 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 450 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 450

# Creating the dataframe
TE_V450 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 300 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 300

# Creating the dataframe
TE_V300 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 150 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 150

# Creating the dataframe
TE_V150 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

### Combining all dataframes and Calculate other variables
TE_fromVWB <- TE_fromVWB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
  # Create other variables
  mutate(D = B*W, # Words produced in the dominant language
         N = (1-B)*W, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= V)) %>%
  # filter out cases where D or N aren't whole numbers
  filter_at(vars(D, N), any_vars(is.wholenumber(.))) 
```

#### Set 2: Calculate other variables based on V, D, and N ####
```{r}
### Generating variables: D and N
## V = 600 ##
V <- 600
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

@@

# Creating the dataframe
TE_V600 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 450 ##
V <- 450
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Creating the dataframe
TE_V450 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 300 ##
V <- 300
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Creating the dataframe
TE_V300 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 150 ##
V <- 150
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Creating the dataframe
TE_V150 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

### Combining all dataframes and Calculate other variables
TE_fromVDN <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # Create other variables
  mutate(W = D+N, # Total word vocabulary of a child
         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>%
  # Remove cases where B = NA
  filter(!is.na(B)) 
```

#### Set 3: Calculate other variables based on V, D, and B ####
```{r}
### Creating variables: D and B
## V = 600 #
V <- 600
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

@@

# Creating the dataframe
TE_V600 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 450 #
V <- 450
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

# Creating the dataframe
TE_V450 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 300 #
V <- 300
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

# Creating the dataframe
TE_V300 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 150 #
V <- 150
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

# Creating the dataframe
TE_V150 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

### Combining all dataframes and Calculate other variables
TE_fromVDB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
  # Create other variables
  mutate(W = D/B, # Total word vocabulary of a child
         N = W-D, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= V)) %>%
  # filter out cases where N isn't whole number
  filter(is.wholenumber(N))
```

#### Set 4: Calculate other variables based on V, N, and B ####
```{r}
### Generating variables: N and B
## V = 600 #
V <- 600
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

@@

# Creating the dataframe
TE_V600 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 450 #
V <- 450
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

# Creating the dataframe
TE_V450 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 300 #
V <- 300
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

# Creating the dataframe
TE_V300 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 150 #
V <- 150
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

# Creating the dataframe
TE_V150 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

### Combining all dataframes and Calculate other variables
TE_fromVNB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
  # Create other variables
  mutate(W = D/(1-B), # Total word vocabulary of a child
         D = W-N, # Words produced in dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= V)) %>%
  # filter out cases where D isn't whole number
  filter(is.wholenumber(D)) %>%
  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.))) %>%
  filter(is.wholenumber(Nu)) 
```

#### Combine the 4 sets ####
```{r}
TE_computation <- TE_fromVWB %>%
  bind_rows(TE_fromVDN, TE_fromVDB, TE_fromVNB) %>%
  # constrain all variables but B (balance) to be integers 
  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
```

# Correlation
```{r}
Correlation <- TE_computation %>%
  cor_matrix()

library(Hmisc)
mycor <- rcorr(as.matrix(TE_computation), type="pearson")
mycor$r
mycor$P

@@

plot_composition_byW <- TE_computation %>%
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N", "C")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  mutate(bin_W = cut_interval(W, 6,
                     labels = c(200, 400, 600, 800, 1000, 1200))) %>%
  group_by(bin_B, vocab_type, bin_W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ bin_W) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
  labs(x = "Balance", 
       y = "Number of concepts", 
       title = "Vocabulary composition across different vocabulary size") 
```

## Compare different vocabulary measures across balance
```{r}
vocab_label <- c("Concept Vocabulary (C)", "Unique Dominant Words (Du)", "Unique Non-dominant Words (Nu)")

plot_words <- TE_computation %>%
  # Convert to long data
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N")) %>%
  mutate(vocab_type = as.factor(vocab_type)) %>%
  mutate(vocab_type = recode(vocab_type, C = "Concept vocabulary (C)", Du = "Unique dominant words (Du)",
         Nu = "Unique non-dominant words (Nu)", TE = "Translation equivalents")) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  #mutate(bin_W = cut_interval(W, 12,
  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
  group_by(bin_B, vocab_type, W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = W, y = n_words, color = bin_B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Total vocabulary Size (W)", 
       y = "Number of words", 
       color = "Balance",
       title = "Word vocabulary size by balance under different vocabulary measures") +
  guides(colour = guide_legend(nrow = 1))
